Combinatorics of the geometry of Wilson loop diagrams II: Grassmann necklaces, dimensions, and denominators

### 2019-08-15 GWLDCII.tex
just a few more things up to p26

We now consider what happens as we move from $V$ to each of the three remaining cases for $W$. We need to consider both $J^{W}_p$ and $J^{W}_r$ for $r \neq p$, since the addition of $p$ can have a knock-on\note{Wouldn't the expression be knock off?} effect on later steps in the algorithm.

@@

\begin{thm}\label{res:alg gives GN}
The sequence of $k$-subsets $(I_1,\dots,I_n)$ obtained by applying Algorithm~\ref{alg:put GN on WLD} to all vertices of an admissible diagram $W$ is a Grassmann necklace.
\end{thm}
\begin{proof}
\note{Sian replaced ``counterclockwise'' with ``clockwise'' in a couple of places in this proof: someone please doublecheck?}
For each $i \in [n]$, let $I_i$ be the set of vertices assigned to the propagators of $W$ by Algorithm~\ref{alg:put GN on WLD} with starting vertex $i$. By Lemma~\ref{GN alg well defined}, we know that $|I_i| = k$ for each $i \in [n]$. We prove that $I_{i+1} \supseteq I_i \backslash \{i\}$ for all $i \in [n]$; this is equivalent to the definition of Grassmann necklace given in Definition~\ref{def:grassmann necklace}.

@@

\textbf{Case I}:  Suppose $p$ has one end on edge $i-1$.  If $p$ is not supported on $i+1$, then in building $I_{i+1}$ we will assign the same propagators as in the construction of $I_i$ from vertices $i+1$ up to $b-1$, that is $I_{i+1} \cap [i+1,b-1] = I_{i} \cap [i+1,b-1]$.  (If $p$ is supported on $i+1$, then $b= i+1$ and the interval $[i+1,b-1]$ is empty.) Furthermore, by Lemma~\ref{vertex cyclic int lem} it must happen that $p$ contributes vertex $b$ to $I_{i+1}$, as otherwise $b$ would never be contributed by $p$. Consequently, in building $I_{i+1}$, when the algorithm reaches vertex $b$ there cannot be any remaining unassigned propagators that are supported on $b$ and clockwise of $p$.  This is also true in when the algorithm constructing $I_i$ reaches $b$, since the same propagators have been assigned beforehand.  Finally, we also note that $m\geq_i b+1$.

@@

\begin{proof}
We have shown that $(I_1,\dots,I_n)$ is a Grassmann necklace; it remains to check that this Grassmann necklace corresponds to the positroid $M(W)$. Recall from the discussion following Definition~\ref{def:grassmann necklace}, and equation \eqref{basesofmatroids} in particular, that it suffices to show: 
\begin{itemize}
\item For each $i \in [n]$, $I_i$ is a basis for $M(W)$.
\item If $J$ is lexicographically smaller than $I_i$ with respect to $<_i$ (for any $i \in [n]$) then $J$ is not a basis for $M(W)$. 
\end{itemize}
The first point follows immediately: the algorithm assigns each propagator to a vertex in the support of that propagator, so $I_i$ is a basis for $M(W)$ by Corollary~\ref{lem basis as perm}.

@@

Now repeat the argument above with $p_s$ in place of $p_r$: if the other end of $p_s$ lands on an edge between $q$ and $j_m$ then we have $q \in \cP_{m-1}$ as above; if not then there exists some $p_t \in Q_{l-3}$ with $p_s$ supported on $I_i(p_t)$ and $q$ on the inside of $p_t$ in the $a_t-1$ order. Thereofore \hlfix{$I_i(p_t) <_i I_i(p_r)$}{$I_i(p_t) <_i I_i(p_s)<_i I_i(p_s)$? Or is the implication clear enough not to be mentioned}\note{Nope: we just need $<_i I_i(p_r)$. $I_i(p_t) <_i I_i(p_r)$ certainly isn't obvious and may not be true}. {\color{violet} May not be necessary, but I think highlighted tex follows from "Now repeat the argument...." \begin{enumerate}\item If $V(p_r) \in [a_r-1, j_m]$ (other end closer to $j_m$) then $q$ inside $p_r$ in $a_r-1$ order. \item Else, $V(p_r) \in [i, a_r+1]$ (Maybe $[i-1, a_r +1]$?). (Main point is that other end closer to $i$) Either way,  admissibility implies that $a_s < a_r$ \item Now we repeat. Either  $V(p_s) \in [a_s-1, j_m]$ then $q$ inside $p_s$ in $a_s-1$ order and we are done \item Else, $V(p_s) \in [i, a_s+1]$ (Maybe $[i-1, a_s +1]$?). Either way, admissibility implies that $a_t < a_s$ and we've already established $a_s < a_r$ \end{enumerate} We don't have to include this argument. It may be totally unnecessary. But I want to make certain that I'm not missing anything.}  Since there are finitely many steps this process can take before reaching $Q_0$, and any $p_* \in Q_0  = \Prop(j_m)$ has $j_m$ in its support by definition, this argument eventually terminates with the construction of a propagator constraining the positioning of $q$ as required. This completes the induction step, and hence the proof.\end{proof}

>>>>

We now consider what happens as we move from $V$ to each of the three remaining cases for $W$. We need to consider both $J^{W}_p$ and $J^{W}_r$ for $r \neq p$, since the addition of $p$ can have a knock-on effect on later steps in the algorithm.

@@

\begin{thm}\label{res:alg gives GN}
The sequence of $k$-subsets $(I_1,\dots,I_n)$ obtained by applying Algorithm~\ref{alg:put GN on WLD} to all vertices of an admissible diagram $W$ is a Grassmann necklace.
\end{thm}
\begin{proof}
\note{Sian replaced ``counterclockwise'' with ``clockwise'' in a couple of places in this proof: someone please doublecheck?  Karen checked and it seems correct with Sian's changes.}
For each $i \in [n]$, let $I_i$ be the set of vertices assigned to the propagators of $W$ by Algorithm~\ref{alg:put GN on WLD} with starting vertex $i$. By Lemma~\ref{GN alg well defined}, we know that $|I_i| = k$ for each $i \in [n]$. We prove that $I_{i+1} \supseteq I_i \backslash \{i\}$ for all $i \in [n]$; this is equivalent to the definition of Grassmann necklace given in Definition~\ref{def:grassmann necklace}.

@@

\textbf{Case I}:  Suppose $p$ has one end on edge $i-1$.  If $p$ is not supported on $i+1$, then in building $I_{i+1}$ we will assign the same propagators as in the construction of $I_i$ from vertices $i+1$ up to $b-1$, that is $I_{i+1} \cap [i+1,b-1] = I_{i} \cap [i+1,b-1]$.  If $p$ is supported on $i+1$, then $b= i+1$ and the interval $[i+1,b-1]$ is empty, so the previous equality remains true, albeit vacuously.  Furthermore, by Lemma~\ref{vertex cyclic int lem} it must happen that $p$ contributes vertex $b$ to $I_{i+1}$, as otherwise $b$ would never be contributed by $p$. Consequently, in building $I_{i+1}$, when the algorithm reaches vertex $b$ there cannot be any remaining unassigned propagators that are supported on $b$ and clockwise of $p$.  This is also true in when the algorithm constructing $I_i$ reaches $b$, since the same propagators have been assigned beforehand.  This also implies that $m\geq_i b+1$.

@@

\begin{proof}
We have shown in the previous theorem that $(I_1,\dots,I_n)$ is a Grassmann necklace; it remains to check that this Grassmann necklace corresponds to the positroid $M(W)$. Recall from the discussion following Definition~\ref{def:grassmann necklace}, and equation \eqref{basesofmatroids} in particular, that it suffices to show: 
\begin{itemize}
\item For each $i \in [n]$, $I_i$ is a basis for $M(W)$.
\item If $J$ is lexicographically smaller than $I_i$ with respect to $<_i$ (for any $i \in [n]$) then $J$ is not a basis for $M(W)$. 
\end{itemize}
The first point follows immediately: the algorithm assigns each propagator to a vertex in the support of that propagator, so $I_i$ is a basis for $M(W)$ by Corollary~\ref{lem basis as perm}.

@@

Now repeat the argument above with $p_s$ in place of $p_r$: if the other end of $p_s$ lands on an edge between $q$ and $j_m$ then we have $q \in \cP_{m-1}$ as above; if not then there exists some $p_t \in Q_{l-3}$ with $p_s$ supported on $I_i(p_t)$ and $q$ on the inside of $p_t$ in the $a_t-1$ order. Therefore \hlfix{$I_i(p_t) <_i I_i(p_r)$}{$I_i(p_t) <_i I_i(p_s)<_i I_i(p_s)$? Or is the implication clear enough not to be mentioned}\note{Nope: we just need $<_i I_i(p_r)$. $I_i(p_t) <_i I_i(p_r)$ certainly isn't obvious and may not be true}. {\color{violet} May not be necessary, but I think highlighted tex follows from "Now repeat the argument...." \begin{enumerate}\item If $V(p_r) \in [a_r-1, j_m]$ (other end closer to $j_m$) then $q$ inside $p_r$ in $a_r-1$ order. \item Else, $V(p_r) \in [i, a_r+1]$ (Maybe $[i-1, a_r +1]$?). (Main point is that other end closer to $i$) Either way,  admissibility implies that $a_s < a_r$ \item Now we repeat. Either  $V(p_s) \in [a_s-1, j_m]$ then $q$ inside $p_s$ in $a_s-1$ order and we are done \item Else, $V(p_s) \in [i, a_s+1]$ (Maybe $[i-1, a_s +1]$?). Either way, admissibility implies that $a_t < a_s$ and we've already established $a_s < a_r$ \end{enumerate} We don't have to include this argument. It may be totally unnecessary. But I want to make certain that I'm not missing anything.}  Since there are finitely many steps this process can take before reaching $Q_0$, and any $p_* \in Q_0  = \Prop(j_m)$ has $j_m$ in its support by definition, this argument eventually terminates with the construction of a propagator constraining the positioning of $q$ as required. This completes the induction step, and hence the proof.\end{proof}
